---
title: "Rjags parameters"
output:
  html_document:
    df_print: paged
---
Adjusting rjags parameters to see how they affect the prob of presence of a particular ASV in the dataset

The jags model feeds from two main sources: 

  * The presence / absence matrix
  
  * The initial parameters for the true positive rate and the false positive rate
  
We can modify the matrix by subsetting the data and doing the detection of an ASV in a month or Site - and later keep all hashes that were true (prob of Occ > 0.8) for a Site or Month.

We can modify the initial parameters by hard-coding a maximum value for the false positive rate; or by picking it up from a left-skewed beta distribution


```{r}
library(tidyverse)
library(vegan)
library(rjags)
library(proxy)
library(here)
```



### Get the abundance table - spread technical replicates

We need to spread the abundance table with as many columns as technical replicates, and with presence/absence data 

```{r}
ASV.nested <- read_rds(here("Cleaning.before.Occ.model"))

ASV.nested %>% 
  select( Step2.tibble) %>% 
  unnest(Step2.tibble) %>% # This stage is a long table
  ungroup() %>%
  mutate(nReads = 1) %>% 
 
  separate(sample, into = c("biol","rep"), sep = "\\.") -> ASV.nested

## To add the absences, get the number of distinct biol

n_distinct(ASV.nested$biol) -> max.biols


```

So the object Pattern of presence has, for each Hash, in how many bottles of water the hash was notfound.appeared-once.twice.thrice.

This again can be nested by regions if you think a global assay is artificially zero-inflated

```{r Pattern of presence}


ASV.nested %>%   
  pivot_wider(values_from = nReads, names_from = rep, values_fill = list (nReads = 0)) %>% 
      mutate (ndetections = `1`+`2`+`3`) %>% 
      group_by(ndetections, Hash) %>% 
      summarise(tot = sum(!is.na(ndetections))) %>% 
  spread(ndetections, tot,fill = 0) %>% 
  mutate(`0` = max.biols - (`1`+`2`+`3`)) -> Hashes.PA.data


Hashes.PA.data %>% 
  unite(repetition.level,`0`,`1`,`2`,`3`, sep = ".") -> Pattern.of.presence 



```


# Functions with the Rjags 

```{r custom functions}

ProbOcc <- function(x, psi, p11, p10, K){
  
  (psi*(p11^x)*(1-p11)^(K-x)) / ((psi*(p11^x)*(1-p11)^(K-x))+(((1-psi)*(p10^x))*((1-p10)^(K-x))))

  } 

jags_for_presence <- function(.x){
  .x %>% 
    transmute( model =  map_dbl (data,
                             function(.y, doprint=FALSE, ni=3000,nt=2,nc=10,nb=1000,myparallel=TRUE){
                               
                               .y %>% ungroup %>% dplyr::select(-biol) -> .y # Reduce the tibble to just the presence/abs matrix
                               
                               jags.inits <- function()(list(psi=runif(1,0.05,0.95),p11=runif(1, 0.01,1),p10=rbeta(1,1,20))) # generates three random starting estimates of psi, p11 and p10
                               jags.data <- list (Y= .y,
                                                  S = nrow(.y),
                                                  K = ncol(.y)) 
                               jags.params <- c("psi","p11","p10")
                               model<-jagsUI::jags(data = jags.data, inits = jags.inits, parameters.to.save= jags.params, 
                                           model.file= "RoyleLink_prior.txt", n.thin= nt, n.chains= nc, 
                                           n.iter= ni, n.burnin = nb, parallel=myparallel)
                               
                               psihat <- model$summary["psi","50%"]
                               p11hat <- model$summary["p11","50%"]
                               p10hat <- model$summary["p10","50%"]    
                               modelSummaries <- model$summary
                               
                               nObs   <- max(rowSums(.y))
                               K <- ncol(.y)
                    
                               model.output <- ProbOcc(nObs, psihat, p11hat, p10hat, K) 
                              
										           return(model.output)
              
                               
                             }))
  
}

jags_unnested <- function(.y, doprint=FALSE, ni=3000,nt=2,nc=10,nb=1000,myparallel=TRUE){
                               
                               .y %>% ungroup %>% dplyr::select(-biol) -> .y # Reduce the tibble to just the presence/abs matrix
                               
                               jags.inits <- function()(list(psi=runif(1,0.05,0.95),p11=runif(1, 0.01,1),p10=rbeta(1,1,20))) # generates three random starting estimates of psi, p11 and p10
                               jags.data <- list (Y= .y,
                                                  S = nrow(.y),
                                                  K = ncol(.y)) 
                               jags.params <- c("psi","p11","p10")
                               model<-jagsUI::jags(data = jags.data, inits = jags.inits, parameters.to.save= jags.params, 
                                           model.file= "RoyleLink_prior.txt", n.thin= nt, n.chains= nc, 
                                           n.iter= ni, n.burnin = nb, parallel=myparallel)
                               
                               psihat <- model$summary["psi","50%"]
                               p11hat <- model$summary["p11","50%"]
                               p10hat <- model$summary["p10","50%"]    
                               modelSummaries <- model$summary
                               
                               nObs   <- max(rowSums(.y))
                               K <- ncol(.y)
                    
                               model.output <- ProbOcc(nObs, psihat, p11hat, p10hat, K) 
                              
										           return(model.output)
              
                               
                             }

model.round.nested <- function(list.hashes,n){
  require(dplyr)
  tibble.out  <- paste("model", n , sep=".")
  tibble.in   <- "data"
  tibble.in   <- rlang::ensym(tibble.in)
  list.hashes %>% 
    mutate(!!tibble.out:= map_dbl(data, jags_unnested))
  #list.hashes <- mutate (list.hashes, !!tibble.out := map (!!tibble.in,  jags_for_presence))
}

```

# Now create the dataset with one entry per unique pattern of presence

So it makes no sense to run the jags for more of one of these combinations

```{r}



ASV.nested %>%   
  pivot_wider(values_from = nReads, names_from = rep, values_fill = list (nReads = 0)) %>% 
  group_by(Hash) %>% 
  nest() -> list.of.Hashes

Pattern.of.presence %>% 
  group_by(repetition.level) %>% 
  slice(1) %>% 
  left_join(list.of.Hashes) %>% 
  select(-Hash) -> ready.to.jags


```

# JAgging

```{r changing p10 to beta 1 20}
sink("RoyleLink_prior.txt")
										cat("model {
										    # Priors
										     psi ~ dunif(0,1)
										    p11 ~ dunif(0.01,1)
										    p10 ~ dbeta(1,20)
										    
										    # Likelihood 
										    for (i in 1:S){
										    z[i] ~ dbern(psi)
										    p[i] <- z[i]*p11 + (1-z[i])*p10
										    for (j in 1:K){
										    Y[i,j] ~ dbern(p[i])
										    }
										    }
										    } ",fill=TRUE)
										sink()
										
										

getwd()


```

Now run, for each combination of patterns, 10 independent chains

```{r}
ready.to.jags %>% slice(1:2) -> res.test
for (i in 1:2){
  test.for.occ.modeling.p10.beta.1.20 <- model.round.nested(list.hashes = ready.to.jags, n = i)
}


(ready.to.jags[1,])


test.for.occ.modeling.p10.beta.1.20 %>% unnest() %>% 
  select(repetition.level, contains("model")) %>% 
write_csv( path = "test.beta.1.20.csv")
# Check the model outputs
test.for.occ.modeling.p10.beta.1.20 <- read_csv("test.beta.1.20.csv") 

test.for.occ.modeling.p10.beta.1.20 %>% 
  gather(contains("model"), key = "Attempt", value = "Output") %>% 
  
  
  ggplot(aes(x = repetition.level, y = Output)) +
  geom_boxplot(outlier.color = "red") + 
  theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank()) +
  geom_hline(yintercept = 0.8, color = "red") +
  ggtitle(label = "p10 rbeta 1,20")

  
```
Once we get this - do a left join with the Pattern of presence and keep only those with acceptable prob of presence - I would say median value <0.8

```{r }

test.for.occ.modeling.p10.beta.1.20 %>% 
  gather(contains("model"), key = "Attempt", value = "Output") %>% 
  group_by(repetition.level) %>% 
  summarise(median = median(Output)) %>% 
  mutate(Keep = case_when(median > 0.8 ~ "Yes",
                          TRUE         ~ "No")) -> pattern.fate
  

Pattern.of.presence %>% 
  left_join(pattern.fate) -> Hashes.fate # In this example ther will be NAs bc I only run the example for a few Patterns

# Youcan add plots and summary stats to see how many unique hashes and nReads are you actually throwing away


```

